---
title: "Chapter 10: Dynamic Regression Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fpp3)
```


__Basic Concept__: a linear model with ARIMA errors. Functionally, a linear regression
is fitted and then an ARIMA is run on the errors of the model. Errors from the 
ARIMA are assumed to be white noise. In R, `ARIMA()` is called with predictor
variables inside the function.


## 10.1 Estimation

Parameters of the model are estimated by minimizing the SSE after the ARIMA has
been performed on the linear regression errors.

If the linear regression errors were minimized instead, valuable information 
would be ignored, statistical tests would be incorrect.

***


## 10.2 Regression w/ ARIMA Errors

`ARIMA()` function fits a regression model with ARIMA errors if exogenous regressors
are included in the call:
`ARIMA(y ~ x + pdq(1,1,0))`
But `pdq()` does not need to be specified as `ARIMA()` can determine it automatically.

__Example__
```{r}
us_change %>%
  pivot_longer(c(Consumption, Income),
               names_to = 'var', values_to = 'value') %>%
  ggplot(aes(x = Quarter, y = value)) +
  geom_line() +
  facet_grid(vars(var), scales = 'free_y') +
  labs(title = "US Consumption and Personal Income",
       y = "Quarterly % Change")

```

It should be clear that this is stationary data, so no differencing will be needed.
```{r}
fit <- us_change %>%
  model(ARIMA(Consumption ~ Income))
report(fit)
```



```{r}
bind_rows(
    `Regression residuals` =
        as_tibble(residuals(fit, type = "regression")),
    `ARIMA residuals` =
        as_tibble(residuals(fit, type = "innovation")),
    .id = "type"
  ) %>%
  mutate(
    type = factor(type, levels=c(
      "Regression residuals", "ARIMA residuals"))
  ) %>%
  ggplot(aes(x = Quarter, y = .resid)) +
  geom_line() +
  facet_grid(vars(type))
```

For Dynamic Regression Models only the ARIMA residuals need to resemble white noise,
and here, they are not differentiated from white noise. The `.innov` are the ARIMA
residuals and `.resid` are the Linear Regression residuals.

```{r}
fit %>% gg_tsresiduals()

augment(fit) %>%
  features(.innov, ljung_box, dof = 5, lag = 8)

```



***


## 10.3 Forecasting

Below we will forecast next eight quarters assuming future percentage changes will
be eequal to the mean for the last 40 years.

__Example__
```{r}
us_change_future <- new_data(us_change, 8) %>%
  mutate(Income = mean(us_change$Income))
forecast(fit, new_data = us_change_future) %>%
  autoplot(us_change) +
  labs(y = "Percentage change")
```

Prediction intervals are narrower than if we had just fit an ARIMA with no covariates
as there is now greater explanation of the data with the Income predictor.

__Caveat__: Prediction intervals from regression models do not take into account
any uncertainty of future values of the predictors and are conditional on the 
assumptions made regarding them, thus the reason for adding 8 quarters of the mean
above.

***


## 10.4 Stochastic and Deterministic Trends

Both Stochastic and Deterministic trend models use ARIMA on the errors. For 
Stochastic I=0; for Deterministic I=1. 


***


## 10.5 Dynamic Harmonic Regression

For data with long seasonal periods (daily data with annual seasonality: 365, 
weekly data: 52, half-hourly data: 48), dynamic regression with Fourier terms
is recommended. Previous models max out at 24.

__Example__
```{r}
aus_cafe <- aus_retail %>%
  filter(
    Industry == "Cafes, restaurants and takeaway food services",
    year(Month) %in% 2004:2018
  ) %>%
  summarise(Turnover = sum(Turnover))

fit <- model(aus_cafe,
  `K = 1` = ARIMA(log(Turnover) ~ fourier(K=1) + PDQ(0,0,0)),
  `K = 2` = ARIMA(log(Turnover) ~ fourier(K=2) + PDQ(0,0,0)),
  `K = 3` = ARIMA(log(Turnover) ~ fourier(K=3) + PDQ(0,0,0)),
  `K = 4` = ARIMA(log(Turnover) ~ fourier(K=4) + PDQ(0,0,0)),
  `K = 5` = ARIMA(log(Turnover) ~ fourier(K=5) + PDQ(0,0,0)),
  `K = 6` = ARIMA(log(Turnover) ~ fourier(K=6) + PDQ(0,0,0))
)

fit %>%
  forecast(h = "2 years") %>%
  autoplot(aus_cafe, level = 95) +
  facet_wrap(vars(.model), ncol = 2) +
  guides(colour = "none", fill = "none", level = "none") +
  geom_label(
    aes(x = yearmonth("2007 Jan"), y = 4250,
        label = paste0("AICc = ", format(AICc))),
    data = glance(fit)
  ) +
  labs(title= "Total monthly eating-out expenditure",
       y="$ billions")
```

For the above:

  * Note the `PDQ()` is declared as the Fourier calls handle seasonality.
  * AICc value is minimized for K=6, which would be the one used.


***


## 10.6 Lagged Predictors

The impact of a predictor may not be immediate. There may be a delay for sales
to catch up to an advertising campaign and sales may be resultingly stronger 
after the campaign concludes.

For situations like this, we can lag the predictors.

Here we will look at TV Ads having an effect on number of quotations an insurance
agency provides.

```{r}
insurance %>%
  pivot_longer(Quotes:TVadverts) %>%
  ggplot(aes(x = Month, y = value)) +
  geom_line() +
  facet_grid(vars(name), scales = "free_y") +
  labs(y = "", title = "Insurance advertising and quotations")
```

Below a study on advertising expenditure for up to four months before the quote data.

```{r}
fit <- insurance %>%
  # Restrict data so models use same fitting period
  mutate(Quotes = c(NA, NA, NA, Quotes[4:40])) %>%
  # Estimate models
  model(
    lag0 = ARIMA(Quotes ~ pdq(d = 0) + TVadverts),
    lag1 = ARIMA(Quotes ~ pdq(d = 0) +
                 TVadverts + lag(TVadverts)),
    lag2 = ARIMA(Quotes ~ pdq(d = 0) +
                 TVadverts + lag(TVadverts) +
                 lag(TVadverts, 2)),
    lag3 = ARIMA(Quotes ~ pdq(d = 0) +
                 TVadverts + lag(TVadverts) +
                 lag(TVadverts, 2) + lag(TVadverts, 3))
  )

glance(fit)
```

The best model looks to be advertising for the current month and one previous month
considered. Then model is re-estimated using all available data.

```{r}
fit_best <- insurance %>%
  model(ARIMA(Quotes ~ pdq(d = 0) +
              TVadverts + lag(TVadverts)))
report(fit_best)
```

In order to forecast, we create new data for the predictor variable with TVadverts 
being an average of 8 for 20 periods.

```{r}
insurance_future <- new_data(insurance, 20) %>%
  mutate(TVadverts = 8)
fit_best %>%
  forecast(insurance_future) %>%
  autoplot(insurance) +
  labs(
    y = "Quotes",
    title = "Forecast quotes with future advertising set to 8"
  )
```

