---
title: "Chapter 04 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(glue)
library(GGally)
library(broom)
```

## 4.6 Exercises

1. Write a function to compute the mean and standard deviation of a time series, and apply it to the PBS data. Plot the series with the highest mean, and the series with the lowest standard deviation.

```{r}
PBS %>%
  as_tibble()%>%
  group_by(ATC1_desc) %>%
  summarize(Avg_Cost = mean(Cost), 
            Std_Cost = sd(Cost)) %>%
  arrange(desc(Avg_Cost))
```

* Series with the highest mean: Cardiovascular System
* Series with the lowest standard deviation: Antiparasitic products, insecticides and repellents


```{r}

PBS %>%
  group_by(ATC1_desc) %>%
  filter(ATC1_desc == 'Cardiovascular system') %>%
  summarize(Total_Cost = sum(Cost)) %>%
  autoplot(Total_Cost) +
  labs(title = "Cardiovascular System Cost")


PBS %>%
  group_by(ATC1_desc) %>%
  filter(ATC1_desc == 'Antiparasitic products, insecticides and repellents') %>%
  summarize(Total_Cost = sum(Cost)) %>%
  autoplot(Total_Cost) +
  labs(title = "Antiparasitic products, insecticides and repellents Cost")
  
```

2. Use GGally::ggpairs() to look at the relationships between the STL-based features for the holiday series in the tourism data. Change seasonal_peak_year and seasonal_trough_year to factors, as shown in Figure 4.3. Which is the peak quarter for holidays in each state?
```{r}
holiday_series <- tourism %>%
  filter(Purpose == 'Holiday')

holiday_series_features <- holiday_series %>%
  features(Trips, feature_set(pkgs = 'feasts'))

holiday_series_features %>%
  select(seasonal_peak_year, seasonal_trough_year, State) %>%
  mutate(
    seasonal_peak_year = seasonal_peak_year +
      4*(seasonal_peak_year == 0),
    seasonal_trough_year = seasonal_trough_year +
      4*(seasonal_trough_year == 0),
    seasonal_peak_year = glue('Q{seasonal_peak_year}'),
    seasonal_trough_year =glue('Q{seasonal_trough_year}'),
  ) %>%
  ggpairs(mapping = aes(color = State)) +
  theme(axis.text.x.bottom = element_text(angle=45, hjust=1))

```

ACT, NSW, South Australia, Tasmania, Victoria, Western Australia peak in Q1
Nothern Territory and Queensland peak in Q3

***

3. Use a feature-based approach to look for outlying series in the PBS data. What is unusual about the series you identify as “outliers.”
```{r}
PBS  
```

```{r}
pbs_features <- PBS %>%
  features(Scripts, feature_set(pkgs = 'feasts'))
```

```{r}

pbs_features <- pbs_features %>%
  drop_na()

pca <- pbs_features %>%
  select(-Concession, -Type, -ATC1, -ATC2) %>%
  prcomp(scale = TRUE) %>%
  augment(pbs_features)

pca %>%
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = ATC1)) +
  geom_point() +
  theme(aspect.ratio = 1)
```

We can see there are outliers here in the bottom right corner.


```{r}
outliers <- pca %>%
  filter(.fittedPC2 < -10) %>%
  select(Concession, Type, ATC1)

outliers %>% 
  left_join(PBS, by = c("Concession", "Type","ATC1")) %>%
  mutate(
    Series = glue("{Concession}","{Type}","{ATC1}", .sep = '\n\n')
  ) %>%
  ggplot(aes(x = Month, y=Cost)) +
  geom_line() +
  facet_grid(Series ~ ., scales = 'free') +
  labs(title = "PCA Outliers for PBS")

```

I have tried the cost and the total scripts variables and I am still unsure as
to why these are outliers. Possible reasons:

* J: high seasonality and increasing variance but always falling back to 0 on the dips
* C: some seasonality and upward trend
* S: large volume increase towards the end of the series

