---
title: "Ch 8 Notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fpp3)
```

# Chapter 8: Exponential Smoothing

__Basic Concept__: Exponential smoothing is a model that weights the averages of
past observations where the weights decay exponentially as observations get older.
In short - newer observations mean more.


## 8.1 Simple Exponential Smoothing (SES)

__SES__: suitable for forecasting data with no clear trend or seasonal pattern

__Example__
```{r}
algeria_economy <- global_economy %>%
  filter(Country == "Algeria")
algeria_economy %>%
  autoplot(Exports) +
  labs(y = "% of GDP", title = "Exports: Algeria")
```
```{r}
# Estimate parameters
fit <- algeria_economy %>%
  model(ETS(Exports ~ error("A") + trend("N") + season("N")))
fc <- fit %>%
  forecast(h = 5)
```

The A's and N's above stand for Additive and NULL, respectively, as this is a 
forecast for a series with no trend or seasonality.

```{r}
fc %>%
  autoplot(algeria_economy) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit)) +
  labs(y="% of GDP", title="Exports: Algeria") +
  guides(colour = "none")
```

Graph shows forecast and one-step-ahead fitted values (orange).

Large prediction interval shows large uncertainty; using the mean value alone 
would be misleading.

***


## 8.2 Methods with Trend

__Holt's Linear Trend__: allowed SES to forecast trending data.


__Example__
```{r}
aus_economy <- global_economy %>%
  filter(Code == "AUS") %>%
  mutate(Pop = Population / 1e6)
autoplot(aus_economy, Pop) +
  labs(y = "Millions", title = "Australian population")
```

Holt's linear trend method is employed by making the `trend()` argument Additive.
Smoothing params for the level, trend, and first fitted values are estimated by 
minimizing SSE.


__Damped Trend Method__: Holt's trend tends to over forecast, especially with
forecasts far into the future. Dampening introduces a param that flattens to 
horizontal as the forecast continues. 

`phi` ($\phi$) is an argument for the dampening amount. At 1, it is the same as using a 
non-dampening method. Usually restricted to values between 0.8 and 9.8.

```{r}
aus_economy %>%
  model(
    `Holt's method` = ETS(Pop ~ error("A") +
                       trend("A") + season("N")),
    `Damped Holt's method` = ETS(Pop ~ error("A") +
                       trend("Ad", phi = 0.95) + season("N"))
  ) %>%
  forecast(h = 15) %>%
  autoplot(aus_economy, level = NULL) +
  labs(title = "Australian population",
       y = "Millions") +
  guides(colour = guide_legend(title = "Forecast"))
```


***

## 8.3 Methods with Seasonality

__Holt-Winter's Method__: contains three smoothing equations (level,
trend, seasonal) each with a smoothing param ($\alpha$, $\beta$, $\gamma$), and 
_m_ to denote the seasonal freq. 

   * Additive Method: preferred when seasonal variations are constant
   
      - Expressed in absolute terms in scale of series
      - series is seasonally adjusted by subtracting the seasonal component
      - Seasonal component adds up to ~0 within 1 year span
      
   * Multiplicative Method: preferred when seasonal variations change proportional
   to the level of the series
   
      - Expressed in percentages
      - Seasonally adjusted by dividing by the seasonal component
      - Seasonal component adds up to _m_ within 1 year span

__Example__
```{r}
aus_holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  summarise(Trips = sum(Trips)/1e3)
fit <- aus_holidays %>%
  model(
    additive = ETS(Trips ~ error("A") + trend("A") +
                                                season("A")),
    multiplicative = ETS(Trips ~ error("M") + trend("A") +
                                                season("M"))
  )
fc <- fit %>% forecast(h = "3 years")
fc %>%
  autoplot(aus_holidays, level = NULL) +
  labs(title="Australian domestic tourism",
       y="Overnight trips (millions)") +
  guides(colour = guide_legend(title = "Forecast"))
```

__Holt Winter's Damped__: can be used with additive and multiplicative methods,
but multiplicative typically provides accurate forecasts for seasonal data

__Example__
Multiplicative model with damped trend "Ad" on daily data with _m_=7.
```{r}
sth_cross_ped <- pedestrian %>%
  filter(Date >= "2016-07-01",
         Sensor == "Southern Cross Station") %>%
  index_by(Date) %>%
  summarise(Count = sum(Count)/1000)
sth_cross_ped %>%
  filter(Date <= "2016-07-31") %>%
  model(
    hw = ETS(Count ~ error("M") + trend("Ad") + season("M"))
  ) %>%
  forecast(h = "2 weeks") %>%
  autoplot(sth_cross_ped %>% filter(Date <= "2016-08-14")) +
  labs(title = "Daily traffic: Southern Cross",
       y="Pedestrians ('000)")
```

Model has identified weekly seasonal pattern and increasing trend 


***


## 8.4 A Taxonomy of Exponential Smoothing Methods

There are a lot of combinations of smoothing methods. 


***


## 8.5 Innovations of State Space Models for Exponential Smoothing

__state space model__: consists of a measurement equation that describes the
observed data and some state equations that describe how the unobserved states
(level, trend, season) change over time.

This statistical models in this section generate forecast intervals via a stochastic
data generation process.

Honestly, all of these models would be figure out automatically with R minimizing
AICc.

***


## 8.6 Model Selection

R automatically uses AICc to choose the best model when using `ETS()`

```{r}
aus_holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  summarise(Trips = sum(Trips)/1e3)
fit <- aus_holidays %>%
  model(ETS(Trips))
report(fit)
```
`report()` outputs both the $\alpha$ and $\gamma$ as well as the estimates for the 
initial states of the params. Small $\gamma$ values indicates seasonal components 
change very little over time.

***

## 8.7 Forecasting with ETS Models

Just like any other model, use `forecast()`.

```{r}
fit %>%
  forecast(h = 8) %>%
  autoplot(aus_holidays)+
  labs(title="Australian domestic tourism",
       y="Overnight trips (millions)")
```


***


