---
title: "Chapter 2 Notes"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fpp3)
library(readr)
```
## 2.1 tsibble objects
Formerly a ```ts()``` object, ```tsibble()``` is the new way to establish a 
timeseries object 

```
y <- tsibble(
  Year = 2015:2019,
  Observation = c(123, 39, 78, 52, 110),
  index = Year
)
```
```tsibble``` objects extend tidy data frames (```tibble`` objects) by introducing
temporal structure.

When observations are more frequent that yearly, a timeclass function must be
used as the index. Below is a monthly `tibble` df:
```z
#> # A tibble: 5 x 2
#>   Month    Observation
#>   <chr>          <dbl>
#> 1 2019 Jan          50
#> 2 2019 Feb          23
#> 3 2019 Mar          34
#> 4 2019 Apr          30
#> 5 2019 May          25
```

In order to convert to `tsibble`, convert the `Month` column from `<chr>` to `<mth>`
using `yearmonth()` and identifying `index` variable with `as_tsibble()`:
```
z %>%
  mutate(Month = yearmonth(Month)) %>%
  as_tsibble(index = Month)
```


<br></br>
**Other Time Class Functions**

Feature|Function
-------|---------
|Annual  | `start:end`|
|Quarterly    | `yearquarter()`|
|Monthly      | `yearmonth()`|
|Weekly       | `yearweek()`|
|Daily        | `as_date()`, `ymd()`|
|Sub-daily    | `as_datetime()`, `ymd_hms()`|


<br></br>
**Working with `tsibble` Objects**

We can use `dplyr` functions on `tsibble` objects.Examples below using the PBS
tsibble containing sales data on pharmaceutical products in Australia
```{r}
PBS
```
Using the `filter()` function to call specific value from column:
```{r}
PBS %>%
  filter(ATC2 == 'A10')
```
Selecting the specific columns we need with `select()`:
```{r}
PBS %>%
  filter(ATC2 == 'A10') %>%
  select(Month, Concession, Type, Cost)

```
`select()` handles columns while `filter()` handles rows

`summarize()` allows you to combine data across keys:
```{r}
PBS %>%
  filter(ATC2 == 'A10') %>%
  select(Month, Concession, Type, Cost) %>%
  summarize(TotalC = sum(Cost))
```
Creating new variables using `mutate()`
```{r}
PBS %>%
  filter(ATC2 == 'A10') %>%
  select(Month, Concession, Type, Cost) %>%
  summarize(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC/1e6)
```
Saving as a tsibble():
```{r}
a10 <- PBS %>%
  filter(ATC2 == 'A10') %>%
  select(Month, Concession, Type, Cost) %>%
  summarize(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC/1e6)
```


<br></br>
**Reading CSVs**
```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv")

# The original CSV has the date variable as individual days and they should be quarters

prison <- prison %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)

prison
```


***


## 2.2 Time Plots

**Example of a time plot**
```{r}
autoplot(a10, Cost) +
  labs(y='$ in millions',
       x='',
      title= 'Australian Antidiabetic Drug Sales')
```

Plot shows:

* clear and increasing trend
* strong seasonal pattern
* increase in variance

Reason behind the shape:

* Government subsidizes in such a way that makes it cost-effective for 
patients to stockpile at the end of the calendar year, which leads to the 
drop at the beginning of each year.



  
***


## 2.3 Time Series Patterns

**Trend**: A long-term increase or decrease in the data.

**Seasonality**: When a pattern occurs in a fixed and known period of time, 
typically a yearly or less (i.e. hourly, weekly, monthly, quarterly).

**Cycle**: Occurs when the data exhibits a pattern not of fixed frequency usually
on a scale > 2 years (i.e. market downturns every 7-10 years ) 


***


## 2.4 Seasonal Plots
A seasonal plot shows the data plotted against each individual "season"
```{r}
a10 %>%
  gg_season(Cost, labels='both') +
  labs(y = '$ in millions',
       x='',
       title = 'Seasonal Plot: Antidiabetic Drug Sales') +
  expand_limits(x = ymd(c('1972-12-28','1973-12-04')))
```

**Multiple Seasonal Periods**
In a case where data has more than one season pattern, use `period` argument.
```{r}
vic_elec %>% gg_season(Demand, period='day') +
  #theme(legend.position = 'none') + #very unclear without the legend
  labs(y='MW', title='Electricity Demand: Victoria')
```
```{r}
vic_elec %>% gg_season(Demand, period='week') +
  labs(y='MW',title='Weekly Electricty Demand: Victoria')
```


***

## 2.5 Seasonal Subseries Plot 

Alternative plot where data from each season is collected in a mini time plot
```{r}
a10 %>%
  gg_subseries(Cost)+
  labs(y = '$ in millions',
       title = 'Australian Antidiabetic Drug Sales'
  )
```

*What the plot shows*:

* Blue horizontal lines indicate means for each month. 
* Seasonal pattern can be seen clearly
* Shows seasonal changes over time


***

## 2.6 