---
title: "Chapter 04 Notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
```

## 4.1 Simple Stats

How to calculate the __mean__:
```{r}
tourism %>%
  features(Trips, list(mean = mean)) %>%
  arrange(mean)
```

How to calculate __quantiles__:
```{r}
tourism %>% features(Trips, quantile)
```

For this the min is labeled 0% and max is labeled 100%. Weird.

***

## 4.2 ACF Features

  * Sum of the first ten squared autocorrelation coefficients is use as a 
  summary for how much autocorrelation there is in a series
  * You can also use autocorrelation on differenced data and seasonally 
  differenced data (more in future chapters)
  
  `feat_acf()` computes some of the autocorrelations above
```{r}
tourism %>% features(Trips, feat_acf)
```

## 4.3 STL Features

* For strongly trended data, seasonally adjusted should have much more variation 
than the remainder. For data with little to no trend, variances should be about the same.
* Calculate trend strength as follows:

$$F_T = max\left(0,1 - \frac{Var(R_t)}{Var(T_t + R_t)} \right)$$

* calculate seasonal strength as follows:

$$F_S = max\left(0,1 - \frac{Var(S_t)}{Var(T_t + R_t)} \right)$$
* A series with strong trend or strong seasonality would be close to 1 where a series with no 
seasonality or trend would be close to 0.

```{r}
tourism %>%
  features(Trips, feat_stl)
```

To identify which series are heavily trended and which are heavily seasonal:

```{r}
tourism %>%
  features(Trips, feat_stl) %>%
  ggplot(aes(x = trend_strength,
             y = seasonal_strength_year,
             col = Purpose)) +
  geom_point() +
  facet_wrap(vars(State))
```

Here we can see holiday travel is, of course, most seasonal while strongest trends
are in Victoria and W Australia

***

## 4.4 Other Features

Just a bullet point list of features of `feasts`
***

## 4.5 Exploring the previous list of bullets

All features of the feasts package can be computed like this:
```{r}
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs = 'feasts'))

tourism_features
```

```{r, fig.width=12, fig.height=12}
library(glue)

tourism_features %>%
  select_at(vars(contains('season'), Purpose)) %>%
  mutate(
    seasonal_peak_year = seasonal_peak_year +
      4*(seasonal_peak_year==0),
    seasonal_trough_year = seasonal_trough_year +
      4*(seasonal_trough_year==0),
    seasonal_peak_year = glue('Q{seasonal_peak_year}'),
    seasonal_trough_year = glue('Q{seasonal_trough_year}')
  ) %>%
  GGally::ggpairs(mapping = aes(colour = Purpose), progress=FALSE)
```

Notable observations:
  * The 3 numerical measures related to seasonality are all positively correlated
  * The bottom left panel and top right panel both show that the most strongly 
  seasonal series is Holiday
  * Bar plots in the bottom row of `seasonal_peak_year` and `seasonal_trough_year`
  columns show that seasonal peaks in Business travel occur most often in Q3 and
  least often in Q1
  

__PCA__
This is where they barely introduce the subject of PCA. Better to learn the 
details from somewhere else.

```{r}
library(broom)

pca <- tourism_features %>%
  select(-State, -Region, -Purpose) %>%
  prcomp(scale = TRUE) %>%
  augment(tourism_features)

pca %>%
  ggplot(aes(x = .fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
```
 YOURE ON 4.5 RIGHT AFTER PCA. EXPLAIN THE ABOVE CHART AND CONTINUE.
 
