---
title: "Chapter 3: Timer Series Decomposition"
output: html_document
---

```{r setup, include=FALSE}
library(fpp3)
```

> _Decomposition definition - A time series is of composed of 3 components: trend, seasonality, and the remainder. Decomposition breaks a time series down into its 3 component parts for further study._ 


## 3.1 Transformations and Adjustments

#### 4 Types

1. __Calendar Adjustments__
    a. Some variation in seasonal data may be due to calendar effects (i.e. 
    some months have fewer days than others and potentially lower sales)
    b. a potential solution to the above would be to compute avg sales per
    day in each month.

2. __Population Adjustments__
    a. Data affected by population changes should likely be adjusted to give per-capita data.
    b. Problem: studying amt of hospital beds in a certain region over time. Consider
    number of beds per 1000 ppl to adjust for population changes in the region.
    
__Example:__

A common transformation of GDP is GDP per-capita
```{r}
global_economy %>%
  filter(Country == 'Australia') %>%
  autoplot(GDP/Population) +
  labs(title= 'GDP per capita', y='$US', x='')
```
    
3. __Inflation Adjustments__

The value of money changes over time, and data associated with currency value should
be adjusted for the inflation of that currency. (i.e. historical house price data
should be in context of the most recent time period in the data, or in today's dollars).

To make such adjustments, a price index is used.

__Example:__
Looking at aggregate annual "newspaper and book" retail turnover from `aus_retail`
and adjusting the data for inflation using Consumer Price Index (CPI) from 
`global_economy` allow us to see the changes over time.
```{r}
print_retail <- aus_retail %>%
  filter(Industry == "Newspaper and book retailing") %>%
  group_by(Industry) %>%
  index_by(Year = year(Month)) %>%
  summarize(Turnover = sum(Turnover))

aus_economy <- global_economy %>%
  filter(Code == 'AUS')

print_retail %>%
  left_join(aus_economy, by = "Year")%>%
  mutate(Adjusted_Turnover = Turnover / CPI * 100) %>%
  pivot_longer(c(Turnover, Adjusted_Turnover),
               values_to = 'Turnover') %>%
  mutate(name = factor(name,
                       levels = c("Turnover", "Adjusted_Turnover"))) %>%
  ggplot(aes(x=Year, y=Turnover)) +
  geom_line() +
  facet_grid(name ~ ., scales = 'free_y')+
  labs(title = "Turnover: Australian Print Media Industry",
       y = '$AU',
       x='')

```

Above we can see that Australia's print industry has been in decline for much
longer than the original time series would suggest.

4. __Mathematical Transformations__

Helpful when a time series has variations that increase or decrease with the level of the series.

  * log transformations
  * power transformations
  * Box Cox transformations - a combination of the above two methods which
  depends on the parameter $\lambda$
      * A good value of $\lambda$ is one which makes the size of the seasonal
      variation about the same across the entire series
      * The `guerrero` feature can be used to choose a value of lambda for you.
      
```{r}
autoplot(aus_production, Gas)+
  labs(y='',
       title='Gas Production with No Tranform')

lambda <- aus_production %>%
  features(Gas, features = guerrero) %>%
  pull(lambda_guerrero)
aus_production %>%
  autoplot(box_cox(Gas, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2))))
```

Above we can see the variance of seasonal ups and downs has been transformed
to a consistent variation.

***

## 3.2 Time Series Components

__Two Types of Decomposition:__

1. __Additive__

  $$y_t= S_t + T_t + R_t$$
  
  A. Where $y_t$ is the data, $S_t$ is the seasonal component, $T_t$ is the trend, and
  $R_t$ is the remainder

  B. Most appropriate if the magnitude of the seasonal fluctuations or varation
  of the trend cycle does not vary with the level of the time series.

2. __Mutliplicative__

  $$y_t = S_t * T_t * R_t$$

  A. Most appropriate when the variation in the seasonal pattern appears to be 
    proportional to the level of the time series (i.e. very common with economic data)
  
  B. An alternative to using a multiplicative series is to transform the data
    (like via Box-Cox in 3.1) until varation is steady over time, then use additive
    decomposition.

  $\log y_t = \log S_t + \log T_t + \log R_t$ is equivalent to the multiplicative equation


__Example:__
```{r}
us_retail_employment <- us_employment %>%
  filter(year(Month) >= 1990, Title == 'Retail Trade') %>%
  select(-Series_ID)

dcmp <- us_retail_employment %>%
  model(stl = STL(Employed))

autoplot(components(dcmp))
```

Trend, Seasonality and the Remainder (White noise) are shown in the bottom three
panels. The gray boxes on the left of each series show the relative scale of each
one of the components as each box is the same size. The large size of the box for 
the remainder component shows that its variation is actually the smallest of the
three.


#### __Seasonally Adjusted Data__

_definition_: the resulting data from when the seasonal component is removed 
from the original time series

Situations where seasonality is not of primary interest:

  * Monthly employment data are usually seasonally adjusted in order to highlight
variation of the underlying state of the economy rather than the seasonal variation.
  * An increase in unemployment do to people leaving school seeking work is seasonal 
  variation while an increase in unemployment due to recession is not.

__Example__

Below shows the seasonal data subtracted from the original series
```{r}
components(dcmp) %>%
  as_tsibble() %>%
  autoplot(Employed, color = 'gray') +
  geom_line(aes(y=season_adjust), color = '#0072B2') +
  labs(y='Persons (thousands)',
       x='',
       title='Total Employment in US Retail')
```

***

## 3.3 Moving Averages

Moving averages are the foundation of estimating the trend portion of decomposition
using the following equation:

$$\hat{T_t} = \frac{1}{m}\sum_{j=-k}^{k}y_t+j$$

where $m = 2k+1$. The estimate of the trend-cycle at time $t$ is obtained by averaging 
values of the time series within $k$ periods of $t$. The average eliminates
some of the randomness in the data. This is termed "$m$-MA", a moving average (MA)
of order $m$.

__Example__
```{r}
global_economy %>%
  filter(Country == 'Australia') %>%
  autoplot(Exports) +
  labs(y='% of GDP',
       x='',
       title='Total Australian Exports')
```

To compute the moving average we would do the following:

```{r}
aus_exports <- global_economy %>%
  filter(Country == 'Australia') %>%
  mutate(
    `5-MA` = slider::slide_dbl(Exports, mean, .before=2, .after=2, .complete=TRUE)
  )
```

Then we plot the original and the MA

```{r}
aus_exports %>%
  autoplot(Exports) +
  geom_line(aes(y = `5-MA`), colour = "#D55E00") +
  labs(y = "% of GDP",
       x='',
       title = "Total Australian exports") +
  guides(colour = guide_legend(title = "series"))
```

(Personal note here: coming form Python, it is unclear to me why the moving average
is calculated based on observations before and after the the current observation.
In Python and in finance, rolling averages are typically calculated only from 
preceding observations. This can obviously be adjusted in the `.before` and `.after` 
arguments above, but it is unclear why a centered MA is being taught.)

Also note, you can take moving averages of moving averages:
```{r}
beer <- aus_production %>%
  filter(year(Quarter) >= 1992) %>%
  select(Quarter, Beer)
beer_ma <- beer %>%
  mutate(
    `4-MA` = slider::slide_dbl(Beer, mean,
                .before = 1, .after = 2, .complete = TRUE),
    `2x4-MA` = slider::slide_dbl(`4-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )

beer_ma

```

__Weighted Moving Averages__

Combinations of moving averages (like the 2x4 above) result in weighted moving 
averages.

A major advantage of the weighted moving average is they yield a smoother estimate,
also, if the MA is not centered, the later observations would have greater weight
and thus more influence on forecasts.

***

## 3.4 Classical Decomposition

The last paragraph says "While classical decomposition is still widely used,
it is not recommended as there are now several much better methods." So I will
move on.

***

## 3.5 Decomp Methods Used by Stats Agencies

__2 Most Popular Methods__

1. X-11
    a. Originated with US Census Bureau
    b. Based on classical decomp but contains extra features:
        i. trend-cycle estimates for end points
        ii. seasonal component allowed to vary over time
        iii. handles trading day variation and holidays
        iv. Methods for both additive and multiplicative decomp
        v. robust to outliers
```{r}
x11_dcmp <- us_retail_employment %>%
  model(x11 = X_13ARIMA_SEATS(Employed ~ x11())) %>%
  components()

autoplot(x11_dcmp) +
  labs(x='', title='Decomposition of Total US Retail Employment Using X-11')
```

Above, the X-11 trend cycle captures the sudden fall due to the 2008 financial
crisis better than STL or classical decomp.It can also be useful to employ seasonal
sub-series plots of the seasonal component to help visualize the variation in the 
seasonal component over time. Here there are only small changes over time:
```{r}
x11_dcmp %>%
  gg_subseries(seasonal)
```


2. SEATS
    a. Stands for Seasonal Extraction ARIMA Time Series (Ch 9 covers ARIMA models)
    b. Developed by Bank of Spain
    c. too complicated to explain apparently
```{r}
seats_dcmp <- us_retail_employment %>%
  model(seats = X_13ARIMA_SEATS(Employed ~ seats())) %>%
  components()

autoplot(seats_dcmp) +
  labs(title = 'Decomposition of Total US Retail Employment Using SEATS')
```

Result is quite similar to X-11.

`X-13ARIMA_SEATS()` function calls the `seasonal` package which has many options
for handling variations.

***


## 3.6 STL Decomposition

1. Stands for "Seasonal and Trend decomposition using Loess"
2. Has advantages over SEATS and X-11:
    a. handles any type of seasonality not just quarterly and yearly
    b. seasonal component may change over time, and rate of change controlled 
    by user
    c. smoothness of trend-cycle can be controlled by user
    d. robust to outliers
3. Has disadvantages:
    a. does not handle trading day or calendar variation automatically
    b. can only be used for additive decomposition
        i. if multiplicative decomp is needed, take log of the data and then
        back-transform components
        ii. any decomps between multiplicative and additive, use Box-Cox
        
__Example__
```{r}
us_retail_employment %>%
  model(STL(Employed ~ trend(window = 7) +
              season(window='periodic'),
            robust = TRUE)) %>%
  components() %>%
  autoplot()
``` 

The two main parameters for `STL()` are trend-cycle window `trend()` and seasonal
window `seasona()`. Both should be odd numbers.

`STL()` by default provides automated values  of seasonal window of 13 and a trend
window of 7. This usually gives a good balance, but in the above case the trend
underfits the data and signal from the 2008 financial crisis shows in the remainder.
(Change the values above to verify.) Selecting the shorter window as above improves
the fit.